Adaptive Waveform Inversion for Transmitted Wavefields

### 2024-09-12 repro/papers/awi/paper.tex
interim

\section{Introduction}
Full waveform inversion (FWI) refers to the inference of mechanical
parameter distribution within a material object  (the earth, a human
body, a manufactured artifact,...) from the measurement of wave motion
at its surface. Over the last several decades, this task has come to
be carried out by minimizing a measure of the difference between computational
prediction of wave data versus actual observed
data. The mechanical parameters to be inferred appear as
coefficients in systems of partial differential equations governing
the wave motion, and discrete representations of these parameters form
the unknown vectors in the optimization problems thus generated.

While widely used in recent years, FWI exhibits a pathology when wave
velocities or their parametric equivalents are amongst the parameters
to be estimated: the excessive sensitivity of predicted waveforms to
changes in velocity leads to oscillations in the difference measures,
especially for the most commonly used square-integral measure. Local
optimization algorithms (relatives of Newton;'s method) are the only practical
methods to generate sequences of approximate optimal
parameters. However the oscillations in the difference measures lead
these sequences to become entrapped at near parameter distributions
far from the optimum, and thus do not extract the maximum information
from the data \cite[]{VirieuxOperto:09}.

@@

While these and other works discuss heuristic justifications for AWI and
many details of implementation, a clear explanation for its effectiveness is 
so far missing from the literature. In this short note, I show that
for active sensing of acoustic materials with unique arrival times of
transmitted waves
(that is, unique rays of geometric asymptotics between sources and
receivers), and known source wavelet, the AWI objective function is an asymptotic (in time
frequency) approximation to the
mean-square time-of-travel misfit, a commonly used objective for
travel time tomography. The error in this approximation is
proportional to the source mean-square wavelength. Thus AWI and travel time fitting
lead to similar material parameter estimates. Since the latter is
widely believed to be a near-optimal method for wave velocity
estimation, this relation explains roughly ``why AWI works'', at
least within this restricted scope.

@@

\subsection{Acoustic Modeling and Waveform Inversion}

Each acoustic source is modeled an isotropic point radiator
with known location $\bx_s\in \bR^3$ and wave form (``wavelet'')
$w(t)$ (idealized as independent of $\bx_s$ for simplicity). These
sources generate acoustic waves traveling through Euclidean space. The
data are samples of these waves at receivers are modeled as isotropic
point sensors, recording (excess) pressure at points
${\bf x}_r \in \bR^3$. The recording takes place over a time interval
$[0,t_{\rm max}]$ (also idealized as independent of $\bx_s$ and
$\bx_r$) . I will refer to the recording interval and the
source-receiver position pairs collectively as the {\em acquisition
  geometry}.

The pressure and velocity fields $p({\bf x},t;{\bf x}_s)$,
${\bf v}({\bf x},t;{\bf x}_s)$ for the source location ${\bf x}_s$
depend on the bulk modulus $\kappa({\bf x})$, buoyancy $\beta({\bf
  x})$ (reciprocal of the density $\rho({\bf x})$), and source wavelet $w(t)$ through the acoustic system
\begin{eqnarray}
  \label{eqn:awe}
 \frac{\partial p}{\partial t} & = &- \kappa \nabla \cdot {\bf v} +
                                    \left(\int^t w \right)
 \delta({\bf x}-{\bf x}_s); \nonumber \\
\frac{\partial {\bf v}}{\partial t} & = & - \beta \nabla p; \\ 
p, {\bf v} & = & 0 \mbox{ for }  t \ll 0.
\end{eqnarray}
The representation of the point source time-dependence as an
indefinite integral of the source wavelet will prove convenient
in expressing the asymptotic approximation of $p$.

For simplicity, I assume that the system \ref{eqn:awe} holds over all of Euclidean
space-time $(\bx,t) \in \bR^3 \times \bR$, and ignore physical
boundaries such as the Earth's surface.
For sake of brevity, define the model vector $m=(\kappa,\rho, w)$. The
{\em modeling operator} $F$ maps the model $m$ to the corresponding data $\{p({\bf x}_r,t;{\bf
  x}_s): 0 \le t \le t_{\rm max}\}$.

For the purposes of this paper, the source wavelet $w$ and the density
$\rho$ will be assumed known. In fact, even if actual acoustic sources are
well-approximated by isotropic point radiators, their wave forms are
usually poorly determined and may reasonably be
included as  parameters to be estimated by inversion of the
data. However for the present discussion, $w$ is regarded as
known.

Acoustic inversion means: given the set of
source-receiver position pairs $\{(\bx_s,\bx_r)\}$, source wavelet
$w(t)$, density $\rho(\bx)$,  and data traces
$\{d(\bx_r,t;\bx_s): 0 \le t \le t_{\rm max}\}$, find a bulk modulus $\kappa$ so that the model
$m=(\kappa,\rho,w)$ satisfies $F[m] \approx d$ in some sense. The
simplest version of FWI concretizes this task by asking for a model
$m$ (that is, a bulk modulus $\kappa$) minimizing the square integral/sum
error over the recording time interval $[0,t_{\rm max}]$ and
source-receiver position pairs:
\begin{equation}
  \label{eqn:fwi}
  J_{\rm FWI}[m;d] =\frac{1}{2} \sum_{{\bf x}_r,{\bf x}_s}
  \int_0^{t_{\rm max}}\,dt\, |F[m]({\bf x}_r,t;{\bf x}_s) - d({\bf x}_r,t;{\bf x}_s)|^2
\end{equation}
The definition \ref{eqn:fwi} makes sense only if the predicted and
observed data traces ($F[m]({\bf x}_r\cdot;{\bf x}_s)$ and $d({\bf
  x}_r,\cdot;{\bf x}_s)$) are square-integrable functions on $[0,t_{\rm max}]$, for each
source-receiver position pair $(\bx_r,\bx_s)$. This property is
assumed for $d$, and follows from additional assumptions on $m$ and the
acqquisition geometry, to be explained below.

It is convenient to introduce the $L^2$ norm symbol $\| \cdot \|$ as
shorthand for the right-hand side of the definition \ref{eqn:fwi}:
\[
  J_{\rm FWI}[m;d]= \frac{1}{2}\|F[m]-d\|^2.
\]
In this notation, the interval $[0,t_{\rm max}]$ and the
source-receiver position pair set $\{(\bx_r,\bx_s)\}$ are implicit. In
the sequel, I will use the $L^2$ norm notation to stand for square
integral/sum over whatever variables make sense in context.

Practical versions of FWI include various types of regularization and
imposition of constraints on $m$, see the references cited in the
Introduction. As mentioned there, local iterative optimization methods
are the only practical approach to minimizing $J_{\rm FWI}$ due to the
computational cost of approximating the forward map $F$ and related
operators, but these tend to stagnate at highly sub-optimal
(``cycle-skipped'') estimates of $m$ unless the initial estimate is of
quite high quality, often available only at considerable cost in data
collection and processing.

\subsection{Source Characteristics}

In applications to seismology, medical imaging, and other wave imaging
technologies, the wavelet $w$ is actually the product of signal
processing aimed to shape the data $F[m]$ into an approximation of
data that would have been predicted using a known wavelet. That is, the wavelet is to some extent {\em
  chosen}. This circumstance justifies several constraints on the
wavelet, consistent with the properties of observed signals: the
wavelet is assumed to be of class $H^1_0(\bR)$, that is, square integrable with
square inttegrable derivative and vanishing for sufficiently large
$t$, and even in $t$.
Introduce asymptotics in this setting by assuming that $w = \wl$ is chosen
from a family $\{\wl:\}$ indexed by the dimensionless parameter
$0 <\lambda <1\}$, and defined by
\begin{equation}
  \label{eqn:wfam}
  \wl(t) = \lambda^{-1/2}w_1(t/\lambda).
\end{equation}
The  ``mother wavelet'' $w_1$ is even and of class $H^1_0(\bR)$, and
all wavelets $\wl$ in this family
inherit these properties from $w_1$, along with its dimensions (energy/time$^2$).

It follows from these assumptions that the Fourier transform $\fwl$ is
real-valued, even, of class $H^1(\bR)$. and its product with frequency $|\omega|$ is also
square-integrable.

@@

\subsection{AWI via Optimization}

@@

Since linear acoustics is time-translation invariant, its solution commutes with time convolution, that is,
\[
 \bar{F}[\bar{m}] = \bar{F}[(\kappa,\rho,,u*w)] = F[m]*u
\]
Introduce the notation $S[m]$ for the convolution operator with kernel $F[m]$:
\begin{equation}
  \label{eqn:sdef}
  F[m]*u \equiv S[m]u
\end{equation}
\cite{Warner:16} assume (implicitly) that there is always an exact
solution of the extended inversion problem. That is, they assume that
for any $m$, there is an extended source $\bar{w}$, or equivalently an
adaptive filter $u[m;d]$ with $\bar{w}=w*u[m;d]$, for which
$S[m]u[m;d]= d$. This proposition is dubious, but is easy to work
around, as will be seen below. Also, \cite{Warner:16} formulate
AWI in terms of the inverse filter, that is, $u$ for which
$u*d = F[m]$, and in fact recommend that definition. The present
discussion uses only the forward filter.

@@

It may not be obvious at first glance why the AWI approach to estimating
$m$, described in the last section, should be any more effective than
FWI. To see why AWI works when FWI does not, in at
least some settings, assume that the spatial dimension is three,
and make use of the geometric asymptotics (ray theory) approximation
to the solution of the wave equation. This approximation is {\em
  local}: it applies when sources and receivers are sufficiently
close. I will discuss its global failure and the consequences for AWI
in the fifth section of this paper.

Since the fields $\kappa$ and $\rho$ have been assumed smooth,
the geometric asymptotics approximation is applicable for solution of the point
radiator problem \ref{eqn:awe}. This approximation expresses the pressure field in terms
of $m$-dependent geometric amplitude $a[m]({\bf x}_s,{\bf x}_r)$ and
travel time $\tau[m]({\bf x}_s,{\bf x}_r)$, and implies that
\begin{equation}
  \label{eqn:hadamard}
  F[m_{\lambda}]({\bf x}_r,t;{\bf x}_s) \approx a[m]({\bf x}_s,{\bf x}_r) \wl(t-\tau[m]({\bf x}_s,{\bf x}_r)). 
\end{equation}
Here and in the notation to follow, $m_{\lambda} =
(\kappa,\rho,\wl)$.

In this approximation, and in other uses of the symbol $\approx$, the
error is dominated by data wavelength, that is, RMS wavelength of the
source wavelet $\wl$, which is proportional to $\lambda$ as mentioned above. In other words, ``$\approx$'' means ``in error by $O(\lambda)$''.

Approximation \ref{eqn:hadamard} follows from a
``progressing wave''  expansion of the fundamental
solution (Green's function) of the wave equation due to Hadamard
\cite[]{Friedlander:75,Qian:JCP24}. For smooth (slowly varying)
$m$, it holds {\em locally}, that is, for $\bx_r$ sufficiently close to $\bx_s$ that a unique ray of
geometric acoustics connect $\bx_s$ to $\bx_r$. This ``single
arrival'' condition fails (in
general) at larger distance. The precise region of validity depends on
$m$. For the remainder of this section, its validity is assumed.

>>>>

\section{Introduction}
Full waveform inversion (FWI) refers to the inference of mechanical
parameter distribution within a material object  (the earth, a human
body, a manufactured artifact,...) from remote measurements of waves
propagating through it. Over the last several decades, this task has come to
be carried out by minimizing a measure of the difference between computational
prediction of wave data versus actual observed
data. The mechanical parameters to be inferred appear as
coefficients in systems of partial differential equations governing
the wave motion, and discrete representations of these parameters form
the unknown vectors in the optimization problems thus generated.

While widely used in recent years, FWI exhibits a pathology when wave
velocities or their parametric equivalents are amongst the parameters
to be estimated: the excessive sensitivity of predicted waveforms to
changes in velocity leads to oscillations in the difference measures,
especially for the most commonly used square-integral measure. Local
optimization algorithms (relatives of Newton;'s method) are the only practical
methods to generate sequences of approximate optimal
parameters. However the oscillations in the difference measures lead
these sequences to stagnate at parameter distributions
far from the optimum, and thus do not extract the maximum information
from the data \cite[]{VirieuxOperto:09}.

@@

While these and other works discuss heuristic justifications for AWI
and many details of implementation, a clear explanation for its
effectiveness is so far missing from the literature. In this short
note, I treat a special case of FWI, the recovery of smooth acoustic
medium parameters from data generated at localized energy sources,
propagated through a fluid, and recorded at receiver locations remote
from the sources. I show that the AWI objective function is an
asymptotic (in time frequency) approximation to the mean-square
time-of-travel misfit, provided that arrival times of transmitted
waves are single, that is, unique rays of geometric asymptotics
connect sources and receivers. The error in this approximation is
proportional to the source mean-square wavelength. Thus AWI and
minimization of mean-square travel time error (travel time tomography)
lead to similar material parameter estimates. Travel time tomography
appears to be much less sensitive to choice of initial model than is
FWI, and in fact is commonly used to construct an initial model for
FWI \cite[]{Bordingetal:87,SirguePratt:04,VirieuxOperto:09}. At least
in the special case treated here, AWI implicitly combines travel time
tomography with FWI.

@@

\subsection{Acoustic Modeling of Transmitted Waves}

Each acoustic source is modeled an isotropic point radiator
with known location $\bx_s\in \bR^3$. These
sources generate acoustic waves traveling through Euclidean space. The
data are samples of these waves at receivers, modeled as isotropic
point sensors, recording pressure at points
${\bf x}_r \in \bR^3$. The recording takes place over a time interval
$[0,t_d]$ (also idealized as independent of $\bx_s$ and
$\bx_r$) . The collection $X_{sr}$ of source-receiver pairs
$(\bx_s,\bx_r)$ for which data is recorded should not intersect the
diagonal - that is, no receiver should be co-located with a source for
which it is active. The set of source locations $X_s = \{\bx_s: (\bx_s,\bx_r) \in X_{sr}
\mbox{ for some } \bx_r \in \bR^3\}$ will also be useful.

In practice, recording takes place over a finite time interval
$[0,t_d]$. For convenience, assume that this interval is the same for
all receivers. Let $\phi \in C_0^{\infty}(0,t_d)$ be a window function
supported in this interval. Additional constraints on $t_d$ and $\phi$
will be be specified below.

The pressure and velocity fields $p({\bf x},t;{\bf x}_s)$,
${\bf v}({\bf x},t;{\bf x}_s)$ for the source location ${\bf x}_s$
depend on the bulk modulus $\kappa({\bf x})$, buoyancy $\beta({\bf
  x})$ (reciprocal of the density $\rho({\bf x})$), and source wave
form $f(t)$ through the acoustic system
\begin{eqnarray}
  \label{eqn:awe}
 \frac{\partial p}{\partial t} & = &- \kappa \nabla \cdot {\bf v} +
                                    f(t) \delta({\bf x}-{\bf x}_s); \nonumber \\
\frac{\partial {\bf v}}{\partial t} & = & - \beta \nabla p; \\ 
p, {\bf v} & = & 0 \mbox{ for }  t \ll 0.
\end{eqnarray}
For simplicity, I assume that the system \ref{eqn:awe} holds over all of Euclidean
space-time $(\bx,t) \in \bR^3 \times \bR$, and ignore physical
boundaries such as the Earth's surface.

As mentioned in the introduction, I will assume that the coefficients
$\rho$ and $\kappa$ are bounded and smooth:
\begin{equation}
  \label{eqn:coeff}
  \log \rho, \log \kappa \in C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3)
\end{equation}
Then the symmetric hyperbolic system \ref{eqn:awe} has a distribution
solution if $f \in \cal{E}'(\bR)$ (for example \cite{Lax:PDENotes}). The Propagation of
Singularities Theorem (for example, \cite{Tay:81}, Theorem VI.2.1)
implies that if $f \in \cal{E}'(\bR)$ and $(\bx_s,\bx_r) \in X_{sr}$,
then the trace $p(\bx_r, \cdot;\bx_s)$ on $\{(\bx_r,t): t \in \bR\}$ is well-defined,
as a distribution vanishing for sufficiently large
negative $t$. Define the {\em modeling operator}
\begin{equation}
  \label{eqn:modop}
  F: (C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3))^2 
  \rightarrow X_{sr} \times {\cal D}'(\bR)
\end{equation}
by
\begin{equation}
  \label{eqn:modopdef}
  F[\kappa,\rho](\bx_r,\bx_s) = \phi p(\bx_r,\cdot;\bx_s).
\end{equation}
For sake of brevity, define the model vector $m=(\kappa,\rho)$.

For the purposes of this paper, the source wavelet $f$ will be assumed
known, and to satisfy
\begin{itemize}
\item[F1. ] $f \in H^1_0(\bR)$,
\item[F2. ] $f$ is odd, $f(t) = -f(-t)$.
\end{itemize}
In fact, even if actual acoustic sources are well-approximated by
isotropic point radiators, their wave forms are only roughly known
{\em a priori} and may reasonably be included as parameters to be
estimated by inversion of the data. On the other hand, in applications
to seismology, medical imaging, and other wave imaging technologies,
the source wave form $f$ is actually the product of signal processing
aimed to shape the data $F[m]$ into an approximation of data that
would have been predicted using a prescribed $f$. That is, the wavelet
is to some extent {\em chosen}. This circumstance justifies the
constraints F1, F2 placed on $f$.

The density $\rho$ will be presumed spatially homogeneous, or at least
to have the same value at all of the source positions $\bx_s\in X_s$:
introduce the {\em source density} $\rho_s > 0$ for which
\begin{equation}
  \label{eqn:rho}
  \rho(\bx_s) = \rho_s\mbox{ for any } \bx_s \in X_s.
\end{equation}
In some settings, eg. the earth's
upper crust, density exhibits substantially less variance than the
bulk modulus, Also, for some acquisition designs, such as marine towed
streamers and some medical ultrasonic technologies, sources are
immersed in water and thus in positions with approximately the same
densities.

In any event, for the present discussion, $f$ is regarded
as known and subject to F1 and F2, and $\rho$ is subject to the constraint \ref{eqn:rho}.

In this setting, acoustic FWI means: given the set of
source-receiver position pairs $X_{sr}$, source wavelet
$f$,  source density $\rho_s$, and data traces
$d:  X_{sr} \rightarrow \cal{D}(\bR)$, find $\kappa$ and $\rho$
(subject to the constraint \ref{eqn:rho}) so that the model
$m=(\kappa,\rho)$ satisfies $F[m] \approx d$ in some sense. If
$d$ and $F[m]$ lie in a normed subspace of $\cal{D}(\bR)$ for each
$(\bx_s,\bx_r) \in X_{sr}$, then the
norm provides a straightforward candidate for this sense of
approximation.

\subsection{Hadamard's Decomposition}
The Hadamard decomposition of the acoustic Green's function provides a
basis for a norm-based formulation of FWI \cite[]{Friedlander:75,Qian:JCP24}. As presented by
\cite{Friedlander:75}, for instance, this construction applies to
second-order wave equations, but the system \ref{eqn:awe} is
equivalent fo a second-order problem: differentiate the first equation
with respect to $t$, eliminate $\bv$ via the second equation, and
rearrange to obtain
\begin{equation}
  \label{eqn:awe2}
  \frac{\rho}{\kappa}\frac{\partial^2 p}{\partial t^2} - \nabla^2 p -
  \nabla \log \rho \cdot \nabla p = \rho \frac{\partial f}{\partial t}
  \delta(\bx-\bx_s)
\end{equation}
The pair $(p,\bv)$ is a distribution solutions of the system
\ref{eqn:awe} if and only if $p$ is a solution of equation
\ref{eqn:awe2} and vis-versa (with $\bv$ constructed from $p$ in the
obvious way, preserving causality).

Hadamard's construction is local, in both space-time and in the set of
acoustic models. A {\em normal neighborhood} of
$\bx_s \in \bR^3$ is an open subset of $\bR^3$ 
within which every point is connected to $\bx_s$ by  a unique ray of
geometric acoustics (a geodesic for the Riemannian metric associated
to the equation \ref{eqn:awe2}) (\cite{Friedlander:75}, p. 16). In the
following, 
$M \subset (C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3))^2$ and $\Omega
\subset \bR^3$ are bounded open sets satisfying 
\begin{itemize}
  \item[G1. ] for every $(\bx_s,\bx_r) \in X_{sr}$, $\bx_s, \bx_r \in \Omega$;
  \item[G2. ] for every source point $\bx_s\in X_{s}$ and $(\kappa,\rho) \in M$, $\Omega$ is a
    normal neighborhood of $\bx_s$.
  \item[G3. ] for every source point $\bx_s \in X_{s};$
    $(\kappa,\rho) \in M$, ray ${\bf X}: [0,t_d] \rightarrow \bR^3$
    with ${\bf X}(0)=\bx_s$, and
    $t \in [0,t_d]$, either $\{{\bf X}(s): 0\le s \le t\} \subset \Omega$
    or $\{{\bf X}(s): t < s < t_d\} \subset \bR^3 \setminus \Omega$.
  \end{itemize}
The last condition means that once a ray leaves $\Omega$ within the
recording time interval $[0,t_d]$, it does not return to $\Omega$
for $t<t_d$.

Note that for any $X_{sr}$ there exist
non-empty open sets $M$ and $\Omega$ satisfying these conditions: any convex open set
$\Omega$ satisfying condition G1 satisfies conditions G2 and G3 for constant
($\bx$-independent) $\kappa$ and $\rho$, and conditions G2 and G3 are open in
the sense of the $C^{\infty}$ and $L^{\infty}$ topologies.

Under the conditions just described, the construction described by
\cite{Friedlander:75} produces for each $m \in M$ functions $\tau[m]$ (travel time) and $a[m]$
(geometric amplitude) on $X_s \times \Omega$ and $b[m]$ (remainder kernel) on $X_s \times \Omega
\times \bR$, satisfying
\begin{itemize}
\item[H1. ] $\tau[m], a[m],$ and $b[m]$ are smooth off the diagonal,
that is, $\tau[m](\bx_s,\cdot) \in C^{\infty}(\Omega \setminus \{\bx_s\})$
and similarly for $a[m],b[m]$;
\item[H2. ] there exists $A>0$ so that for any $(\kappa,\rho) \in M$
  and $(\bx_s,\bx_r) \in X_{sr}$,
  $|\log a[m](\bx_s,\bx_r)| < A$;
\item[H3. ] for any $t>0$, there is $B\ge 0$ so that for
$(\bx_s,\bx_r) \in X_{sr}, -\infty < s \le t$, and $(\kappa,\rho) \in
M$, $|b[m](\bx_s,\bx_r,s)| +
|\partial_tb[m](\bx_s,\bx_r,s)| \le B$.
\end{itemize}

The Hadamard
decomposition of the Green's function $G$, i.e. the solution of the
equation \ref{eqn:awe2} for $f = H/\rho_s$, is (suppressing the
dependence on $m \in M$ from the notation)::
\begin{equation}
  \label{eqn:green}
  G(\bx_s,\bx,t) = a(\bx_s,\bx)\delta(t - \tau(\bx_s,\bx)) +
  b(\bx_s,\bx,t)H(t - \tau(\bx_s,\bx))
\end{equation}
Under conditions G1 - G3 above, this decomposition is the restriction of
the Greeen's function to $\Omega \times (-\infty,t_d]$. Accordingly,
\[
   F[m](\bx_s,\bx_r)(t) =
   \phi(t)a(\bx_s,\bx_r)\rho_s\frac{df}{dt}(t-\tau(\bx_s,\bx_r))
 \]
\begin{equation}
  \label{eqn:fwdgreen}
 + \phi(t)\rho_s\int
  \,ds \, \frac{df}{dt}(t-s) b(\bx_s,\bx_r,s)H(s-\tau(\bx_s,\bx_r))
\end{equation}

\subsection{FWI via Least Squares}
It follows from the Hadamard decomposition 
(equation \ref{eqn:fwdgreen}) that $F[m]$ is
square-integrable. Indeed, for each $(\bx_s,\bx_r) \in X_{sr}$, the first term on the RHS of equation \ref{eqn:fwdgreen} is clearly
square-integrable, since $f \in H^1_0(\bR)$.
An $L^2$ bound on the second (remainder) term follows from the
pointwise bounds on $b$ described in the last section and Young's
inequality. In fact, there exists $C>0$ so that for $(\kappa,\rho) \in
M$,
\begin{equation}
  \label{eqn:fwdnorm}
  \|F[m]\|_d \le C
  \left\|\frac{df}{dt}\right\|_{L^2(\bR)}.
\end{equation}
in which the left-hand side uses the abbreviation
\[
  \| \cdot \|_d = \| \cdot \|_{L^2(X_{sr} \times \bR)}
\]

This observation justfies posing FWI as a least-squares problem: given $d \in
L^2(X_{sr}\times \bR)$, $f \in H^1_0(\bR)$, and $\rho_s>0$, find $m = (\kappa,\rho)$ to minimize
\begin{equation}
  \label{eqn:fwi}
  J_{\rm FWI}[m;d]= \frac{1}{2}\|F[m]-d\|_d^2.
\end{equation}
Practical versions of least-squares FWI include various types of regularization and
imposition of constraints on $m$, see the references cited in the
Introduction.
%As mentioned there, local iterative optimization methods
%are the only practical approach to minimizing $J_{\rm FWI}$ due to the
%computational cost of approximating the forward map $F$ and related
%operators, but these tend to stagnate at highly sub-optimal
%(``cycle-skipped'') estimates of $m$ unless the initial estimate is of
%quite high quality, often available only at considerable cost in data
%collection and processing.

\subsection{Source Characteristics and Asymptotics}

Note that in the representation \ref{eqn:fwdgreen}, only the
derivative $df/dt$ appears. For convenience, introduce
\begin{equation}
  \label{eqn:wavelet}
  w = \rho_s \frac{df}{dt}
\end{equation}
From the assumed
properties of $f$, $w \in L^2(\bR)$ and has compact support, has zero
mean, and is
even in $t$. Since $f$ with the properties F1 - F3 corresponds
biuniquely with $w$ with the properties just mentioned, I will also
refer to $w$ as ``the wavelet'', and view $F$ as parametrized by $w$.

Introduce asymptotics in this setting by assuming that $w = \wl$ is chosen
from a family $\{\wl:\}$ indexed by the dimensionless parameter
$0 <\lambda \le 1\}$, and defined by
\begin{equation}
  \label{eqn:wfam}
  \wl(t) = \lambda^{-1/2}w_1(t/\lambda).
\end{equation}
The  ``mother wavelet'' $w_1$ is square-integrable of compact support
even, and mean-zero. Define $t_w$ to be the diameter of the support:
\begin{equation}
  \label{eqn:wdiam}
  \mbox{supp }w_1 = [-t_w,t_w].
\end{equation}
All wavelets $\wl$ in this family
inherit these properties from $w_1$, along with its dimensions
(energy/time$^2$). It follows from these assumptions that the Fourier transform $\fwl$ is
real-valued, even, of class $H^1(\bR)$. The corresponding source waveform $f_{\lambda}$ is
given by
\begin{equation}
  \label{eqn:flambda}
  f_{\lambda}(t) = \int_{-\infty}^t\wl =
  \lambda^{1/2}f_1\left(\frac{t}{\lambda}\right).
\end{equation}
Note that
\begin{equation}
  \label{eqn:lambdanorms}
  \|\wl\|_{L^2(\bR)} = \|w_1\|_{L^2(\bR)},\,\|f_{\lambda}\|_{L^2(\bR)}
  = \lambda \|f_1\|_{L^2(\bR)}.
\end{equation}

%and its product with frequency $|\omega|$ is also
%square-integrable.

@@

The modeling operator with wavelet $\wl$ is denoted $\Fl$. From the
decomposition \ref{eqn:fwdgreen}, $\Fl$ is the sum of two terms, the second of which is
\[
   \phi(t)\rho_s\int
   \,ds \, \frac{df_{\lambda}}{dt}(t-s) b(\bx_s,\bx_r,s)H(s-\tau(\bx_s,\bx_r))
\]
\[
  =\phi(t)\rho_s
  f_{\lambda}(t-\tau(\bx_s,\bx_r))b(\bx_s,\bx_r,\tau(\bx_s,\bx_r))
\]
\[
    -
    \phi(t)\rho_s\int
   \,ds \, f_{\lambda}(t-s) (\partial_t
   b(\bx_s,\bx_r,s))H(s-\tau(\bx_s,\bx_r))
\]
The $L^2$ norms of the summands are bounded by a multiple of
$\lambda$, uniform over $(\bx_s,\bx_r) \in X_{sr}, m \in M$, thanks to
the norm scaling \ref{eqn:lambdanorms}, Young's inequality, and conclusion H3.

I will write ``$+ \Ol$'' to mean an  error uniformly bounded in $L^2$ by
a multiple of $\lambda$ over $m \in M$. Thus
\begin{equation}
  \label{eqn:fwdapprox}
  \Fl[m](\bx_s,\bx_r) = \phi a(\bx_s,\bx_r)\wl(t-\tau(\bx_s,\bx_r)) +
  \Ol
\end{equation}
\subsection{AWI via Optimization}

@@

The adaptive filters $u$ are presumed to vanish for $|t| > t_u$, with
the same maximum time $t_u>0$ for all $(\bx_s,\bx_r) \in X_{sr}$, and
to be square integrable: $u \in L^2(X_{sr}\times [-t_u,t_u])$.

Since linear acoustics is time-translation invariant, its solution commutes with time convolution.
%\[
% \bar{F}[\bar{m}] = \bar{F}[(\kappa,\rho,,u*w)] = F[m]*u
% \]
Denote by $\bar{p}(\bx,t;\bx_s,\bx_r), \bar{\bv}(\bx,t;\bx_s,\bx_r)$
the solution of the acoustic system \ref{eqn:awe} with $w(t)$ in the
right-hand side of the first equation replaced by
$\bar{w}(\bx_r,\cdot;\bx_s) = u({\bf  x}_r,\cdot;{\bf
  x}_s)*w$. Then
\[
  \bar{p}(\bx,\cdot;\bx_s,\bx_r) = u({\bf  x}_r,\cdot;{\bf
    x}_s)*p(\bx,\cdot;\bx_s)
\]
Introduce the notation $S[m]$ for the convolution operator with kernel $F[m]$:
\begin{equation}
  \label{eqn:sdef}
  S[m]u(\bx_r,t;\bx_s) \equiv \phi(t)\bar{p}(\bx_r,t;\bx_s,\bx_r).
\end{equation}

\cite{Warner:16} assume (implicitly) that there is always an exact
solution of the extended inversion problem. That is, they assume that
for any $m$, there is an extended source $\bar{w}$, or equivalently an
adaptive filter $u[m;d]$ with $\bar{w}=w*u[m;d]$, for which
$S[m]u[m;d]= d$. This proposition is dubious, but is easy to work
around, as will be seen below - so for the moment accept it.

Also, \cite{Warner:16} formulate
AWI in terms of the inverse filter, that is, $u$ for which
$\phi u*d = F[m]$, and in fact recommend that definition. The present
discussion uses only the forward filter.

@@

It may not be obvious at first glance why the AWI approach to estimating
$m$, described in the last section, should be any more effective than
FWI. Under the smoothness and ray geometry assumptions made above (equation
\ref{eqn:coeff}, conditions G1 - G3) hold, an asymptotic relation
between AWI and travel time tomography emerges, justifying the notion
that AWI is more likely to succeed than is FWI. This relation is based
on the Hadamard decomposition, and therefore is {\em
  local}: it applies when sources and receivers are sufficiently
close to lie in a common normal neighborhood of the source locations,
for all coefficient pairs $(\kappa,\rho)$ in the search set $M$ of the
inverse problem. The relation to travel time tomography is asymptotic
in wavelength/frequency, and conveniently expressed via the wavelet
family $\{\wl\}$ and the associated family of modeling operators
$\Fl$. With these assumptions, the AWI objective function $J_{\rm AWI}[m;d_{\lambda}$
for noise-free data
\begin{equation}
  \label{eqn:noisefree}
  d_{\lambda}=\Fl[m^*]
\end{equation}
is $\lambda$-asymptotic to the mean-square difference between the
source-receiver travel times for $m$ and $m*$.

This relation breaks down if the sources and receivers fail to belong
to a normal neighborhood (that is, if multiple ray paths exist between
sources and receivers), as explained in the fifth section of this paper.

With these assumptions,
the decomposition \ref{eqn:fwdgreen} is applicable for solution of the point
radiator problem \ref{eqn:awe} with extended source
$u(\bx_r,\cdot;\bx_s)*\wl$, for each source-receiver pair
$(\bx_s,\bx_r)\in X_{sr}$. Write $\bar{p}_{\lambda}$ for the pressure
component of the solution, and specialize the definition
\ref{eqn:sdef} to
\begin{equation}
  \label{eqn:sdeflam}
  S_{\lambda}[m]u(\bx_r,t;\bx_s) \equiv \phi(t)\bar{p}_{\lambda}(\bx_r,t;\bx_s,\bx_r).
\end{equation}
Following the reasoning leading to
equation\ref{eqn:fwdapprox}, obtain
\begin{equation}
  \label{eqn:slamapprox}
  S_{\lambda}[m]u({\bf x}_r,t;{\bf x}_s) = \phi(t)a[m]({\bf x}_s,{\bf
    x}_r) (u(\bx_r,\cdot;\bx_s)*\wl)(t-\tau[m]({\bf x}_s,{\bf x}_r)) + O(\lambda).
\end{equation}

I promised additional constraints on the data time window $[0,t_d]$
and cutoff function $\phi$. These are posed in terms of the extreme
traveltimes over $m \in M$. Since all derivatives of the log wave velocity $\log c =
(\log \kappa - \log \rho )/2$ are bounded in the closure of $\Omega$,
uniformly over $m \in M$, 
the ray trace construction of travel time implies that
$0 < \tau_{\rm min} < \tau_{\rm max}$ exist for which $\tau_{\rm min}
\le \tau[m](\bx_s,\bx_r) \le \tau_{\rm max}$ for $(\bx_s,\bx_r) \in
X_{sr}$, $m \in M$.
The additional constraints on $t_d$ and $\phi$,
\begin{itemize}
\item[T1. ] $t_d >  \tau_{\rm max} + t_w + t_u $,
\item[T2. ] $0 < \tau_{\rm min} - t_w - t_u$,
\item[T3. ] $\phi \equiv 1$ on $[\tau_{\rm min} - t_w - t_u,
\tau_{\rm max} + t_w + t_u]$.
\end{itemize}
That is, the support of the leading term in \ref{eqn:slamapprox} is
contained in $\{t: \phi(t) = 1\}$. These assumptions are physically natural, as
they imply that the recording time interval $[0,t_d]$ contains the
support of the extended source data, up to an error vanishing as
$\lambda \rightarrow 0$.

From T3, equation \ref{eqn:slamapprox} can be re-written as
\begin{equation}
  \label{eqn:slambis}
  S_{\lambda}[m]u({\bf x}_r,t;{\bf x}_s) = a[m]({\bf x}_s,{\bf
    x}_r) (u(\bx_r,\cdot;\bx_s)*\wl)(t-\tau[m]({\bf x}_s,{\bf x}_r)) + O(\lambda).
\end{equation}

Similarly, from the definition \ref{eqn:noisefree} and equation
\ref{eqn:fwdgreen},
\begin{equation}
  \label{eqn:noisefreeapprox}
  d_{\lambda}(\bx_r,t;\bx_s) = a[m^*](\bx_s,\bx_r) \wl(t-\tau[m^*](\bx_s,\bx_r)) + O(\lambda).
  \end{equation}
